---
title: "Practical Machine Learning - Prediction Project Writeout"
output: html_document
---

Introduction
------------
The goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to perform analysis to determine what activity an individual has performed.

Load the necessary libraries
----------------------------
```{r, echo=TRUE}
library(caret)
library(randomForest)
library(foreach)
```

Load the training data
----------------------
```{r, echo=TRUE}
training_data <- read.csv("pml-training.csv", header = TRUE)
```

Clean the data
--------------
1. Replace the #DIV/0! as NA
2. Clean the missing data
3. Drop some useless variables that are not used
4. Convert factor variables to numerics

```{r, echo=TRUE}
training_data <- read.csv("pml-training.csv", header = TRUE, na.strings=c("#DIV/0!"))
checkMissData <- (colSums(is.na(training_data))) / (dim(training_data)[1])
names(which(checkMissData > 0.5 ))
training_data [,names(which(checkMissData > 0.5 ))] <- list(NULL)
training_data[ , c("user_name", "X", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp" )] <- list(NULL)
classe <- training_data$classe
training_data$classe <- NULL
for (i in colnames(training_data)) { training_data[,i] <- as.numeric(training_data[,i]) }
training_data$classe <- classe
```

Plot the variable to analyse if skew is detected before creating partitions.

```{r, echo=TRUE}
plot(training_data$classe)
```

Create data partitions
----------------------
Set the seed to 1234 to make it reproducible.
We will use 70% for training and 30% for testing.

```{r, echo=TRUE}
set.seed(1234)
inTrain <- createDataPartition(training_data$classe, p=0.7, list=FALSE)
training <- training_data[inTrain,]
testing <- training_data[-inTrain,]
```

Building a model
----------------
By trial and error, ntree of 200 is used.
```{r, echo=TRUE}
RF_model <- randomForest( classe ~ ., data=training, importance=TRUE, ntree=200)
RF_model
```

Confusion matrix indicates 0.33% which is quite accurate. We will use it on the testing data.

```{r, echo=TRUE}
predictions <- predict( RF_model, newdata=testing[,-(dim(testing)[2]) ])

confusionMatrix(predictions, testing$classe )
```

Confusion matrix indicates 0.9978% which is very accurate. We can also see that the values of sensitivity and specificity have good values; thus the model is good. Predictions look accurate too.

Run the prediction using the model built
----------------------------------------
The same steps were done to the testing data. 
The results of the prediction are then shown.

```{r, echo=TRUE}
testing_data <- read.csv("pml-testing.csv", header = TRUE, na.strings=c("#DIV/0!"))
testing_data[,names(which(checkMissData > 0.5 ))] <- list(NULL)
testing_data[ ,c("user_name", "X", "raw_timestamp_part_1", "raw_timestamp_part_2", "cvtd_timestamp" )] <- list(NULL)
for (i in colnames(testing_data)) { testing_data[,i] <- as.numeric(testing_data[,i]) }

predictions_new <- predict(RF_model, newdata = testing_data[,-(dim(testing_data)[2])])

predictions_new
```

Write the prediction into files.

```{r, echo=TRUE}

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_", i ,".txt")
    write.table(x[i], file = filename, quote = FALSE,
                row.names = FALSE, col.names = FALSE)
  }
}

pml_write_files(predictions_new)
```
